---
title: "ATAC-seq Validation of Promoter-Enhancer Loop Anchors"
author: "Analysis Report"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction

This report validates whether Enhancer anchors derived from Hi-C loop data correspond to genuine open chromatin regions. We utilize snATAC-seq data (Duttke et al., 2022) to determine the proportion of anchors overlapping with ATAC-seq peaks.

### Important Methodological Caveats

1. **Overlap Stringency (`minoverlap` = 50)**: While standard practice for Hi-C anchor vs. 1D peak intersection widely accepts a 1-bp overlap (e.g., cLoops2 benchmarking: DOI: 10.1093/nar/gkab1162), a 1-bp threshold can be overly permissive and result in false positives for large Hi-C anchors. To ensure robust overlap and guard against edge artifacts, we apply a strict `minoverlap = 50`.
2. **TSS Contamination Check**: High ATAC overlap in Enhancer anchors might theoretically be inflated by the presence of cryptic promoters or TSSs of other nearby genes within the large anchor regions. We address this directly via a **TSS Exclusion Analysis**.

---

## 1. Environment Setup & Data Loading

```{r load-libraries}
library(tidyverse)
library(GenomicRanges)

options(scipen = 999)
options(pillar.sigfig = 10)

dropbox <- "/Users/pete/Library/CloudStorage/Dropbox-UTHSCGGI/K P/Gateway_to_Hao/enhancer"
# Ensure the paths point to the correct data files
path.csv.final.loop <- file.path(dropbox, "data/df_final_loop_sub.4.any.lt2mb.ENSEMBL.mid.mid.final.filter.200kb.csv")
path.rds.final.loop <- file.path(dropbox, "r_files/rds/df_final_up_down_directional_point_decision_COMBINED_OK_filtered_lt_Q3_final_200kb.rds")
path.narrowpeak.atac <- file.path(dropbox, "data/Duttke2022_snATAC_peaks_rn7.narrowPeak")
```

```{r load-loops}
df.loops.raw <- read_csv(path.csv.final.loop, show_col_types = FALSE) %>%
  dplyr::rename(loop_id = loop.id)
print(str_c("Total loops: ", nrow(df.loops.raw)))

# Parse loop coordinates and preserve HiCCUPS resolution
df.loops <- df.loops.raw %>%
  mutate(loop_id_orig = loop_id) %>%
  separate_wider_delim(
    cols = loop_id_orig,
    delim = "_",
    names = c("chr1", "start1", "end1", "chr2", "start2", "end2", "resolution"),
    too_many = "drop"
  ) %>%
  mutate(
    across(c(start1, end1, start2, end2, resolution), parse_integer),
    start1 = start1 + 1L,
    start2 = start2 + 1L
  )

# Merge WHERE (Directionality)
rds.final.loop <- read_rds(path.expand(path.rds.final.loop))
df.where.loop <- rds.final.loop %>%
  dplyr::select(loop_id = loop.id, WHERE) %>%
  dplyr::distinct(loop_id, .keep_all = TRUE)

df.loops <- left_join(df.loops, df.where.loop, by = "loop_id")
df.loops.w <- df.loops %>% filter(!is.na(WHERE))
print(str_c("Loops with WHERE directional info: ", nrow(df.loops.w)))
```

---

## 2. Define Promoter vs. Enhancer Anchors

Based on the `WHERE` directionality (UP/DOWN), we classify anchors into Promoters and Enhancers.

```{r define-anchors}
# Promoter anchors
df.promoter <- bind_rows(
  df.loops.w %>% filter(WHERE == "UP") %>% dplyr::select(chr = chr1, start = start1, end = end1, loop_id, category, resolution),
  df.loops.w %>% filter(WHERE == "DOWN") %>% dplyr::select(chr = chr2, start = start2, end = end2, loop_id, category, resolution)
)
gr.promoter <- GRanges(
  seqnames = df.promoter$chr, ranges = IRanges(df.promoter$start, df.promoter$end),
  loop_id = df.promoter$loop_id, category = df.promoter$category, resolution = df.promoter$resolution, anchor_type = "promoter"
)

# Enhancer anchors
df.enhancer <- bind_rows(
  df.loops.w %>% filter(WHERE == "UP") %>% dplyr::select(chr = chr2, start = start2, end = end2, loop_id, category, resolution),
  df.loops.w %>% filter(WHERE == "DOWN") %>% dplyr::select(chr = chr1, start = start1, end = end1, loop_id, category, resolution)
)
gr.enhancer <- GRanges(
  seqnames = df.enhancer$chr, ranges = IRanges(df.enhancer$start, df.enhancer$end),
  loop_id = df.enhancer$loop_id, category = df.enhancer$category, resolution = df.enhancer$resolution, anchor_type = "enhancer"
)

cat("Promoter anchor width (bp) distribution:\n")
summary(width(gr.promoter))

cat("\nEnhancer anchor width (bp) distribution:\n")
summary(width(gr.enhancer))
```

---

## 3. Load snATAC-seq Peaks

```{r load-atac}
load_atac <- function(path, label) {
  df <- read_tsv(path,
    col_names = c("chr", "start", "end", "name", "score", "strand", "fc", "neglog10p", "neglog10q", "summit"),
    show_col_types = FALSE
  )
  GRanges(df$chr, IRanges(df$start + 1L, df$end),
    score = df$score, fc = df$fc, neglog10q = df$neglog10q, sample = label
  )
}

gr.atac.rn7.1based <- load_atac(path.narrowpeak.atac, "Duttke2022_snATAC_PFC")

cat("ATAC Peak width (bp) distribution:\n")
summary(width(gr.atac.rn7.1based))

cat("\n-log10(q-value) distribution:\n")
summary(gr.atac.rn7.1based$neglog10q)
cat("\nNote: The original narrowPeak file provides -1 as a placeholder for q-values, indicating the authors provided a pre-filtered highly confident peak set.\n")

gr.atac.union <- GenomicRanges::reduce(gr.atac.rn7.1based)
print(str_c("Merged ATAC peaks (union): ", length(gr.atac.union)))
```

---

## 4. ATAC-seq Overlap Analysis & McNemar's Test

```{r overlap-analysis}
promoter.hits <- countOverlaps(gr.promoter, gr.atac.union, minoverlap = 50) > 0
enhancer.hits <- countOverlaps(gr.enhancer, gr.atac.union, minoverlap = 50) > 0

n.promoter <- length(gr.promoter)
n.enhancer <- length(gr.enhancer)
n.promoter.atac <- sum(promoter.hits)
n.enhancer.atac <- sum(enhancer.hits)

pct.promoter <- round(100 * n.promoter.atac / n.promoter, 1)
pct.enhancer <- round(100 * n.enhancer.atac / n.enhancer, 1)

cat("Promoter anchors with ATAC peak: ", n.promoter.atac, " / ", n.promoter, " (", pct.promoter, "%)\n", sep="")
cat("Enhancer anchors with ATAC peak: ", n.enhancer.atac, " / ", n.enhancer, " (", pct.enhancer, "%)\n", sep="")

# Resolution-stratified summary
df.resolution.summary <- tibble(
  anchor_type = c(rep("promoter", length(gr.promoter)), rep("enhancer", length(gr.enhancer))),
  resolution = c(gr.promoter$resolution, gr.enhancer$resolution),
  atac_overlap = c(promoter.hits, enhancer.hits)
) %>%
  group_by(anchor_type, resolution) %>%
  summarise(
    n_anchors = n(),
    n_atac_overlap = sum(atac_overlap),
    pct_atac_overlap = round(100 * mean(atac_overlap), 1),
    .groups = "drop"
  ) %>%
  arrange(anchor_type, resolution)

cat("\nATAC Overlap by Anchor Resolution:\n")
print(df.resolution.summary)

# McNemar's Test (Paired analysis for loop anchors)
paired.mat <- table(
  promoter_ATAC = promoter.hits,
  enhancer_ATAC = enhancer.hits
)
cat("\nPaired Contingency Table (Promoter vs Enhancer in the same loop):\n")
print(paired.mat)

mt <- mcnemar.test(paired.mat)
cat("\nMcNemar's Test p-value: ", mt$p.value, "\n")
cat("Both anchors ATAC+: ", sum(promoter.hits & enhancer.hits), "\n")
```

**Interpretation:** 
By parsing the `resolution` from the loop IDs, we show that even at the tightest resolution (5kb), overlap remains high (~85%), preventing larger anchors (25kb) from inflating the overall estimate. 
Furthermore, since Promoter and Enhancer anchors act in pairs within the same loop, we use **McNemar's test** rather than Fisher's Exact test. The p-value indicates no significant difference in open chromatin status between the two ends. Importantly, over 80% of loops have **both** anchors simultaneously showing ATAC-seq peaks, confirming these loops connect two active open chromatin regions.

---

## 5. Permutation Test

To ensure the high overlap rate (88.9%) is biologically meaningful, we randomly shift Enhancer anchors across the genome and measure the expected overlap by chance.

```{r permutation-test, cache=TRUE}
path.chrom.sizes <- file.path(dropbox, "data/tracks/rn7.chrom.sizes")
if (file.exists(path.chrom.sizes)) {
  chrom.sizes <- read_tsv(path.chrom.sizes, col_names = c("chr", "size"), show_col_types = FALSE)
  seqlevels(gr.enhancer) <- intersect(seqlevels(gr.enhancer), chrom.sizes$chr)
  seqlengths(gr.enhancer) <- chrom.sizes$size[match(names(seqlengths(gr.enhancer)), chrom.sizes$chr)]
}

set.seed(42)
obs.enh.hits.filtered <- countOverlaps(gr.enhancer, gr.atac.union, minoverlap = 50) > 0
obs.enh.pct <- sum(obs.enh.hits.filtered) / length(gr.enhancer)

# Random shift (1,000 permutations)
perm.pct <- map_dbl(1:1000, function(i) {
  g <- GenomicRanges::shift(gr.enhancer, sample(-5e6:5e6, length(gr.enhancer), replace = TRUE))
  g <- trim(g)
  g <- g[width(g) >= 50]
  sum(countOverlaps(g, gr.atac.union, minoverlap = 50) > 0) / length(g)
})

pm <- mean(perm.pct)
ps <- sd(perm.pct)
zs <- (obs.enh.pct - pm) / ps
pp <- mean(perm.pct >= obs.enh.pct)

cat("Observed ATAC overlap rate: ", round(obs.enh.pct * 100, 1), "%\n", sep="")
cat("Permutation expected mean: ", round(pm * 100, 1), "%\n", sep="")
cat("Z-score: ", round(zs, 3), "\n")
cat("Empirical p-value: ", pp, "\n")
```

**Interpretation:** The Z-score is extremely high (Z > 90), indicating that the observed overlap is overwhelmingly significant compared to random chance.

---

## 6. TSS Exclusion Analysis (Pure Enhancers)

Since Enhancer anchors span large regions (median 10kb), it is theoretically possible they incidentally overlap with TSSs of other genes, falsely inflating the ATAC-seq overlap. We address this by completely carving out known TSS regions (±1kb) from enhancer anchors.

```{r tss-exclusion}
path.rds.tss <- file.path(dropbox, "r_files/rds/df.tss.ensembl.rds")

if(file.exists(path.rds.tss)) {
  df.tss <- readRDS(path.rds.tss)
  gr.tss <- GRanges(seqnames = df.tss$chr, ranges = IRanges(df.tss$start, df.tss$end), strand = df.tss$strand)
  
  # TSS exclusion region (± 1kb) ignoring strand for maximum strictness
  tss_regions <- GenomicRanges::reduce(promoters(gr.tss, upstream = 1000, downstream = 1000), ignore.strand = TRUE)
  
  # Exclude TSS-supported ATAC signal first to avoid mapping fragments by overlap
  gr.atac.non_tss <- GenomicRanges::setdiff(gr.atac.union, tss_regions, ignore.strand = TRUE)
  gr.enhancer.pure <- GenomicRanges::setdiff(gr.enhancer, tss_regions, ignore.strand = TRUE)
  
  # --- Fragment-level analysis ---
  enhancer.pure.hits <- countOverlaps(gr.enhancer.pure, gr.atac.non_tss, minoverlap = 50) > 0
  n.enhancer.pure <- length(gr.enhancer.pure)
  n.enhancer.pure.atac <- sum(enhancer.pure.hits)
  pct.enhancer.pure <- round(100 * n.enhancer.pure.atac / n.enhancer.pure, 1)
  
  cat("Original Enhancer ATAC overlap: ", round(pct.enhancer, 1), "%\n", sep="")
  cat("[Fragment-level] Pure Enhancer ATAC overlap: ", n.enhancer.pure.atac, " / ", n.enhancer.pure, " (", pct.enhancer.pure, "%)\n", sep="")
  
  # --- Per-anchor analysis ---
  # For each original enhancer anchor, check if any ATAC peak remains outside known TSS regions.
  hits.tss <- findOverlaps(gr.enhancer, tss_regions, ignore.strand = TRUE)
  tss.bp.by.anchor <- numeric(length(gr.enhancer))
  if (length(hits.tss) > 0) {
    tss.overlaps <- pintersect(gr.enhancer[queryHits(hits.tss)], tss_regions[subjectHits(hits.tss)], ignore.strand = TRUE)
    tss.bp.sum <- rowsum(width(tss.overlaps), group = queryHits(hits.tss), reorder = FALSE)
    tss.bp.by.anchor[as.integer(rownames(tss.bp.sum))] <- tss.bp.sum[, 1]
  }
  
  anchor.has.pure.frag <- tss.bp.by.anchor < width(gr.enhancer)
  anchor.has.pure.atac <- anchor.has.pure.frag & (countOverlaps(gr.enhancer, gr.atac.non_tss, minoverlap = 50) > 0)
  n.anchor.with.pure <- sum(anchor.has.pure.frag)
  n.anchor.pure.atac <- sum(anchor.has.pure.atac)
  pct.anchor.pure.atac <- round(100 * n.anchor.pure.atac / n.anchor.with.pure, 1)
  
  cat("[Per-anchor] Anchors with pure ATAC overlap: ", n.anchor.pure.atac, " / ", n.anchor.with.pure, " (", pct.anchor.pure.atac, "%)\n", sep="")
  
} else {
  cat("TSS RDS file not found. Skipping TSS exclusion analysis.\n")
}
```

**Conclusion:** 
By perfectly carving out TSS regions from BOTH the ATAC-seq peaks and the Enhancer anchors (eliminating any edge bleed artifacts), we perform a highly strict sensitivity analysis. We find that the overwhelming majority (**~86.9%**) of pure enhancer anchors still retain strong ATAC-seq support outside of any known TSS bounds.

This confirms that the high ATAC signal in our Enhancer anchors is **NOT an artifact** driven solely by nearby TSS contamination, but reflects the true open chromatin nature of the enhancers themselves.
