---
author: ''
subtitle: <h1>Palmer Lab - UCSD</h1>
affiliation: Palmer Lab- UCSD
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    theme: united
    highlight: tango
    toc: yes
    toc_float: yes
    number_sections: yes

---

```{r,echo=F,eval=T,message=F}
title_var <- basename(getwd())
```

---
title: `r title_var`
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cool functions ive used, echo=FALSE,eval=TRUE,message=FALSE,include=F}

specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))

```

```{r libraries, echo=FALSE,eval=TRUE,message=FALSE,include=F}


library(DT)
library(htmltools)
library(readxl)
library(httr)
library(RJSONIO)
library(openxlsx)
library(purrr)
library(plyr)
library(kableExtra)
library(data.table)
#library(RPostgreSQL)

```


```{r traits and covariates, echo=FALSE,eval=TRUE,message=FALSE,include=F}


load("data/raw_data/traits.RData")


```

******
# **Phenotype information**

<strong><em>PI</strong></em>: `r PI`
<strong><em>Domain</strong></em>: `r domain`
<strong><em>Sub-domain</strong></em>: `r sub_domain`



******
# **Trait descriptions**
Data dictionary absent


```{r trait descriptions, echo=FALSE,eval=F,message=FALSE,include=T}

desc=read.csv("data/raw_data/dictionary.csv",header = T,stringsAsFactors = F)




datatable(desc[,c("trait","description")],rownames = FALSE,extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    columnDefs = list(list(className = 'dt-center', targets="_all"))
  ))





```







******
# **Raw trait distribution**

```{r raw trait distribution,echo=FALSE,eval=F,message=FALSE,include=T,warning=F,results='asis'}

plots <- list.files(path = "pheno_processing_summary/plots/traits", pattern = ".png", recursive = F,full.names=T)

plots<-unique (grep(paste(paste0(traits,".png"),collapse="|"), plots, value=TRUE))


plot_names<-basename(plots)

plot_names<-gsub(".png","",plot_names)


traits_plots<-gsub("\\."," with ",plot_names)

traits_plots<-gsub("_"," ",plot_names)




   for(i in 1:length(plots)){

    #cat(paste0("# ",traits_plots[i],"{#",anchors[i],"}   \n"))
    cat("  \n")
    cat("  \n##",  traits_plots[i], "  \n")
    cat(paste0("![",traits_plots[i],"](", plots[i],")\  "))
    cat("  \n")
    cat('\\pagebreak  ')
    #cat(paste0(traits_plots[i]))
    }





```





******
# **Phenotype processing details **
Quantile normalize separately for males and females then regress out the effect of age / bodyweight / cohort / box number if significant and explains more than 2% of variance**
<br>

+ Following table lists the percent variance explained by age
<br>
<br>

```{r age, echo=FALSE,eval=TRUE,message=FALSE,include=T,comment=NA}
if (file.exists("pheno_processing_summary/covs/age_regressed.RData")){
  load("pheno_processing_summary/covs/age_regressed.RData")
 # need_to_regress=combined_box_cohort_regressed
  summary_lm_age_to_regress$r.squared<-format(round(summary_lm_age_to_regress$r.squared,2),nsmall=2)
  summary_lm_age_to_regress$p.value<-format(summary_lm_age_to_regress$p.value,scientific = T)





if(length(summary_lm_age_to_regress)>0){

summary_lm_age_to_regress$trait<-gsub("qnormed_","",summary_lm_age_to_regress$trait)
datatable(summary_lm_age_to_regress[,c("trait","age_cov","r.squared","p.value","center")],colnames = c("trait","covariate","percent variance explained","p.value","project_name"),rownames=F,extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    columnDefs = list(list(className = 'dt-center', targets="_all"))
  ))
}else{
cat("No covariates were significant and explain more than 2% of variance")

}
}else{
cat("No Covariates were tested")
}


```



+ Following table lists the percent variance explained by experiment box
<br>
<br>

```{r experiment specific covariates, echo=FALSE,eval=TRUE,message=FALSE,include=T,comment=NA}
if (file.exists("pheno_processing_summary/covs/experiment_specific_covariates_regressed.csv")){
 experiment_specific<-read.csv("pheno_processing_summary/covs/experiment_specific_covariates_regressed.csv")
 experiment_specific$pve<-format(round(experiment_specific$pve,2),nsmall=2)
  experiment_specific$p_val<-format(experiment_specific$p_val,scientific = T)





if(length(experiment_specific)>0){

experiment_specific$trait<-gsub("qnormed_","",experiment_specific$trait)
datatable(experiment_specific[,c("trait","cov","cnames","pve","p_val","project_name")],colnames = c("trait","covariate","name","percent variance explained","p.value","project_name"),rownames=F,extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
    columnDefs = list(list(className = 'dt-center', targets="_all"))
  ))
}else{
cat("No covariates were significant and explain more than 2% of variance")

}
}else{
cat("No Covariates were tested")
}


```







******
# **SNP Heritability Estimates _h^2^_**

```{r SNP heritability estimates, echo=FALSE,eval=TRUE,message=FALSE,include=T}
heritability_summary=read.csv(paste0("results/snp_h2_with_N.csv"),header=T,stringsAsFactors = F)
heritability_summary$snp_heritability<-format(round(heritability_summary$snp_heritability,3),nsmall=3)
heritability_summary$SE<-format(round(heritability_summary$SE,3),nsmall=3)
#heritability_summary$Pval<-format(round(heritability_summary$Pval,3),nsmall=3)

heritability_summary<-heritability_summary[,c("trait","N","snp_heritability","SE")]
heritability_summary<-heritability_summary[which(heritability_summary$trait %in% traits),]

heritability_summary %>%
  datatable(
    rownames = FALSE,
    filter = list(position = 'top', clear = FALSE),
    extensions = c("SearchPanes", "Select", "Buttons"),
    options = list(
      dom = "Btip",
      buttons = list('copy',"searchPanes",list(
        extend = 'collection',
        buttons = c('csv', 'excel', 'pdf'),
        text = 'Download'
      )),
          language = list(searchPanes = list(collapse = "Filter Rows")),
      columnDefs = list(
        list(searchPanes = list(show = FALSE), targets = 0),
        list(searchPanes = list(controls = FALSE), targets = 0:1),
        list(className = "dt-center", targets = 0:1)
      )
    )
  )



```


```{r read in QTL table,message=F,warning=F,eval=T,include=T,echo=F}
#new=read.csv("results/snp_h2_all_final_traits_with_N_fullName.csv",header = T,stringsAsFactors = F)
qtls<-read.csv(paste0("results/qtls/table3_final.csv"),header=T,stringsAsFactors = F)



qtls$trait_ref_name<-paste0(gsub("_"," ",qtls$trait)," ",qtls$topsnp)

qtls$trait_id<-seq(1:nrow(qtls))
qtls$trait_rmd<-gsub("_"," ",qtls$trait)
num_of_qtls<-count(qtls,vars="trait_rmd")
qtls$LD_interval_size_mb<-qtls$LD_interval_size_bp/1000000
qtls$LD_interval_size_mb_formatted<-paste0(prettyNum(qtls$LD_interval_size_mb,digits=2)," Mb")

qtls$lz<-paste0(qtls$trait_N,"_",qtls$chr,"_",qtls$pos)


qtls$genes<-paste0('https://rest.rgd.mcw.edu/rgdws/genes/',gsub("chr","",qtls$chr),'/',qtls$LD_interval_start,'/',qtls$LD_interval_stop,"/360")
#Fetch RGD ID and create HTTP links
for(i in 1:nrow(qtls)){


qtls_genes<-fromJSON(qtls$genes[i])
temp<-map_dbl(qtls_genes,"rgdId")
gene_names<-map_chr(qtls_genes,"symbol")
urls<-vector(mode="character", length=length(temp))
gene_urls<-vector(mode="character",length=length(temp))
for(j in 1:length(temp)){

  urls[j]<-paste0("https://rgd.mcw.edu/rgdweb/report/gene/main.html?id=",temp[j])
  gene_urls[j] <- paste0("[", gene_names[j], "](", urls[j], ")")
}

qtls$gene_urls[i]<-paste(gene_urls,collapse = " ")


}

#remove extra digits from QTL table, only keep 3
#af se and beta
qtls$af<-specify_decimal(qtls$af, 3)
qtls$se<-specify_decimal(qtls$se, 3)
qtls$beta<-specify_decimal(qtls$beta, 3)
qtls$topsnp_log10P<-specify_decimal(qtls$topsnp_log10P, 3)




```

```{r variant annotation table, message=F,warning=F,eval=T,include=T,echo=F}
f <- paste0("results/QTLs/",expname,"_snp_annotations_final.csv")
if (file.exists(f)){
ann<-read.csv(paste0("results/QTLs/",expname,"_snp_annotations_final.csv"),header=T,stringsAsFactors=F)
#ann<-ann[unique(grep(paste(nicsa_traits,collapse="|"),ann$trait)),]
ann$trait<-sapply(strsplit(ann$trait_snp,split="_chr"),`[`,1)
snps<-sapply(strsplit(ann$trait_snp,split="_chr"),`[`,2)
ann$trait_pos<-sapply(strsplit(snps,split="_"),`[`,2)



ann$trait_ref_name<-paste0(gsub("_"," ",ann$trait)," ",ann$chr,":",ann$trait_pos)



ann$gene_urls<-NA

#mtcars$mpg <- paste0('<a href="',urls,'">',mtcars$mpg ,"</a>")

ann$gene_urls <- paste0('<a href="',ann$RGD_link,'">',ann$snp_gene,"</a>")

}else{

   ann=data.frame(dummy=character(0))


}

```


```{r prep phewas tables,message=F,warning=F,eval=T,include=T,echo=F}


load("results/phewas/annotated_phewas_tables.RData")




```


```{r read in eQTL table,message=F,warning=F,eval=T,include=T,echo=F,cache=TRUE}
eqtl<-read.csv(paste0("results/eqtl/eqtl_results_annotated.csv"),header=T,stringsAsFactors = F)
eqtl=unique(eqtl)


eqtl$trait_ref_name<-paste0(gsub("_"," ",eqtl$trait)," ",eqtl$trait_topsnp)
eqtl$trait_topsnp<-paste0(eqtl$trait,"_",eqtl$trait_topsnp)

#replace regions with full names





#colnames(eqtl)[which(colnames(eqtl) %in% "proportion_snps_above_threshold")]<-"eqtl_snps_above_threshold"

#gene URLs
#need to skip if No RGD entry
eqtl$gene_urls<-NA



for(i in 1:nrow(eqtl)){

eqtl$gene_urls[i]<- paste0("[", eqtl$gene_name[i], "](", eqtl$RGD_link[i], ")")


}

eqtl$RGD_link[which(is.na(eqtl$RGD_link))]<-"No RGD entry"

eqtl$pubmed_link<-NA


#colnames(eqtl)[grep("\\bvariant_id\\b",colnames(eqtl))]<-"eqtl_topsnp"

#colnames(eqtl)[grep("\\bpval_nominal\\b",colnames(eqtl))]<-"eqtl_topsnp_pval"
eqtl$eqtl_topsnp_log10P <- -log10(eqtl$eqtl_topsnp_pval)
eqtl$eqtl_topsnp_log10P <- format(round(eqtl$eqtl_topsnp_log10P,3),nsmall=3)


colnames(eqtl)[grep("\\brsquare\\b",colnames(eqtl))]<-"r2"
eqtl$r2 <- format(round(eqtl$r2,3),nsmall=3)
#eqtl$dprime <- format(round(eqtl$dprime,3),nsmall=3)





for(i in 1:nrow(eqtl)){
if(!is.na(eqtl$pubmed_url[i])){


eqtl$pubmed_link[i]<- paste0("[", eqtl$gene_name[i], "](", eqtl$pubmed_url[i], ")")
}

}





#gwas catalog link
#https://www.ebi.ac.uk/gwas/genes/CSMD1

eqtl$gwas_C_link<-NA


for(i in 1:nrow(eqtl)){

if(!is.na(eqtl$gwas_catalog_url[i])){
eqtl$gwas_C_link[i]<- paste0("[", eqtl$gene_name[i], "](", eqtl$gwas_catalog_url[i], ")")
}

}






eqtl$gene_long_name_formatted<-paste0("_",eqtl$gene_long_name,"_")


eqtl$alliance_link<-NA
#https://www.alliancegenome.org/search?mode=table&q=csmd1

for(i in 1:nrow(eqtl)){

if(!is.na(eqtl$human_entrez_id[i])){
  url<-paste0("https://www.alliancegenome.org/search?mode=table&q=",tolower(eqtl$gene_name[i]))
eqtl$alliance_link[i]<- paste0("[", eqtl$gene_name[i], "](", url, ")")
}



}




```

******
# **Summary of QTLs**

\hfill\break

```{r summary of QTLs, include=T,message=F,echo=F,warning=F}
qtls_summary_table<-qtls

rownames(qtls_summary_table)<-NULL
datatable(qtls_summary_table[,c("trait","topsnp","topsnp_log10P","LD_interval_size_mb_formatted","num_of_genes")],colnames = c("trait","topSNP","topSNP -log10P","size of interval","Number of genes in interval"),rownames = FALSE,
    filter = list(position = 'top', clear = FALSE),
    extensions = c("SearchPanes", "Select", "Buttons"),
    options = list(
      dom = "Btip",
      buttons = list('copy',"searchPanes",list(
        extend = 'collection',
        buttons = c('csv', 'excel', 'pdf'),
        text = 'Download'
      )),
          language = list(searchPanes = list(collapse = "Filter Rows")),
      columnDefs = list(
        list(searchPanes = list(show = FALSE), targets = 0),
        list(searchPanes = list(controls = FALSE), targets = 0:1),
        list(className = "dt-center", targets = 0:1)
      )
    )
  )




```

******
# **Manhattan plots**


```{r test plotting GWAS,message=F,warning=F,eval=T,include=T,results = 'asis',echo=F}


plots <- list.files(path = "results/manhattan_plots/option1", pattern = ".jpg", recursive = F,full.names=T)

plots<-unique (grep(paste(paste0(qtls$trait,".png"),collapse="|"), plots, value=TRUE))

plot_names<-basename(plots)



plot_names<-gsub(".jpg","",plot_names)


traits_plots<-gsub("\\."," with ",plot_names)

traits_plots<-gsub("_"," ",plot_names)






   for(i in 1:length(plots)){

    #cat(paste0("# ",traits_plots[i],"{#",anchors[i],"}   \n"))
    cat("  \n")
    cat("  \n##",  traits_plots[i], "  \n")
    #cat("  \n Number of QTLs: ",  num_of_qtls$freq[which(num_of_qtls$trait_rmd %in% traits_plots)], "  \n  ")
    cat("  \n    ")

    count_chrs<-count(qtls[which(qtls$trait_rmd %in% traits_plots[i]),],vars="chr")

    colnames(count_chrs)[2]<-"number of qtls"
    cat("  \n")
    cat(paste0("![",traits_plots[i],"](", plots[i],")\  "))
    cat("  \n")
    if(nrow(count_chrs)>0){
    qtl_info_manhattan<-qtls[which(qtls$trait_rmd %in% traits_plots[i]),c("chr","topsnp","LD_interval_size_mb_formatted","topsnp_log10P")]
    count_chrs<-merge(count_chrs,qtl_info_manhattan,by="chr")

    colnames(count_chrs)[3]<-"topSNP"
    colnames(count_chrs)[4]<-"size of interval"
    colnames(count_chrs)[5]<-"topSNP -log10P"
    print(kable(count_chrs) %>% kable_styling(latex_options =c("striped")))
    }
    cat('\\pagebreak  ')
    #cat(paste0(traits_plots[i]))
    }



```


******
# **Regional Association plots**



```{r dynamic chunks,message=F,warning=F,eval=T,include=T,results = 'asis',echo=F,cache=T}


plots <- list.files(path = "results/locuszoom_plots/", pattern = ".png", recursive = F,full.names=T)




qtls$lz_png_name<-NA



for(i in 1:nrow(qtls)){
qtls$lz_png_name[i]<-plots[which(plots %like% qtls$lz[i])]


}

plots<-plots[which(plots %in% qtls$lz_png_name)]
plots<-plots[match(qtls$lz_png_name,plots)]


for(i in 1:nrow(qtls)){
cat("<br>")
cat("  \n")
#prints the QTL name aka heading
cat("  \n##",  qtls$trait_ref_name[i], "  \n")
#prints size of interval and no of genes in interval
cat("  \n Size of interval : ",  prettyNum(qtls$LD_interval_size_bp[i],big.mark = ","), "bp  \n  ")
cat("  \n Number of genes in interval: ",  qtls$num_of_genes[i], "  \n  ")



cat("<br>")
cat("  \n")

#prints locuszoom plots
cat(paste0("![",qtls$trait_id[i],"](", plots[i],")\  "))

cat("  \n")

#prints RGD links for genes in QTL below the locuszoom plot
cat("  \n  RGD link for genes in interval: ",  qtls$gene_urls[i], "  \n  ")

#prints af, se, beta and SDP of founders
print(htmltools::tagList(datatable(qtls[i,c("topsnp","af","se","beta","allele1","allele2",'ACI','BN','BUF','F344','M520','MR','WN','WKY')],rownames = F,options = list(lengthChange = FALSE))))

#cat("<br>")
cat("  \n")
cat("  \n")
cat("  \n### Putatively causal coding variants:   ",  qtls$trait_ref_name[i], "  \n")

if(nrow(ann)>0){


ann_test<-ann[which(ann$trait_ref_name %in% qtls$trait_ref_name[i]),]


if(nrow(ann_test)==0){
cat("  \n HIGH or MODERATE impact variants absent \n  ")
cat("<br>")
cat("  \n")
}else{
print(htmltools::tagList(datatable(ann_test[,c('chr','pos','ref','alt','snp_effect','snp_impact','snp_gene',"gene_urls",'snp_feature','snp_biotype','snp_cdna_pos','snp_hgvs_c','snp_hgvs_p',"r2_with_trait_topsnp","dprime_with_trait_topsnp")],colnames=c("chr","pos","ref","alt","effect","impact","gene","RGD link","feature","biotype","cDNA position","SNP change","Amino acid change","r2 with trait topsnp","dprime"),rownames=F,escape = F,options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
))))
cat("<br>")
cat("  \n")

}



}




else{
cat("  \n HIGH or MODERATE impact variants absent \n  ")
cat("<br>")
cat("  \n")
}

#eQTL info
cat("  \n")
cat("  \n### eQTL info: ",  qtls$trait_ref_name[i], "  \n")
eqtl_test<-eqtl[which(eqtl$trait_ref_name %in% qtls$trait_ref_name[i]),]


if(is.na(eqtl_test$ensembl_gene[1])){
cat("  \n No cis EQTLs detected \n  ")
cat('\\pagebreak  ')
}else{

#print trait topsnp and log10P
cat("  \n Trait topSNP : ", qtls$topsnp[i], "  \n  ")
cat("  \n Trait topSNP -log10P : ",qtls$topsnp_log10P[i], "  \n  ")


#split eqtl test by gene name
eqtl_list<-split(eqtl_test,eqtl_test$gene_name)


for(j in 1:length(eqtl_list)){

cat("  \n#### ",  names(eqtl_list)[j] , "  \n")
#pubmed_link
#gwas_C_link

temp_gene<-eqtl_list[[j]]
cat("  \n Ensembl gene name : ", temp_gene$ensembl_gene[1], "  \n  ")
if(temp_gene$RGD_link[1]=="No RGD entry") {
cat("  \n  _RGD link:_ _No RGD entry_  \n  ")
}else{
cat("  \n<strong>",temp_gene$gene_long_name_formatted[1],"</strong>  \n  ")
cat("  \n  _RGD link:_ ",  temp_gene$gene_urls[1], "  \n  ")
if(!is.na(temp_gene$gwas_C_link[1])){
cat("  \n  _Human GWAS Catalog link:_ ",  temp_gene$gwas_C_link[1], "  \n  ")
}else{
cat("  \n  _Human GWAS Catalog link:_ _No Human GWAS Catalog entry_  \n  ")
}
if(!is.na(temp_gene$pubmed_link[1])){
cat("  \n  _Pubmed link:_ ",  temp_gene$pubmed_link[1], "  \n  ")
}else{
cat("  \n  _Pubmed link:_ _NA_  \n  ")
}
if(!is.na(temp_gene$alliance_link[1])){
cat("  \n  _Alliance of Genome Resources link:_ ",  temp_gene$alliance_link[1], "  \n  ")
}else{
cat("  \n  _Alliance of Genome Resources link:_ _NA_  \n  ")
}
print(htmltools::tagList(datatable(temp_gene[,c("tissue","eqtl_topsnp","eqtl_topsnp_log10P","r2")],colnames = c("tissue","eqtl topSNP","eqtl topSNP -log10P","r2 with trait topSNP"),rownames = F,options = list(dom = 'Bfrtip',buttons = c('copy', 'csv', 'excel', 'pdf', 'print')))))
#cat("<br>")
cat("  \n")
}

}


}

#PheWAS links
cat("  \n")
cat("  \n### PheWAS: P-values for other phenotypes at trait topSNP : \n ",  qtls$trait_ref_name[i], "  \n")



if(nrow(table1_unfiltered)>0){

    phewas_unfilt<-table1_unfiltered[which(table1_unfiltered$trait_ref_name %in% qtls$trait_ref_name[i]),]
    if(nrow(phewas_unfilt)>0){
    rownames(phewas_unfilt)<-NULL
    phewas_unfilt$trait<-gsub("_"," ",phewas_unfilt$trait)

    print(htmltools::tagList(datatable(phewas_unfilt[,c("trait","p_score","log10P","trait_description")],colnames=c("Trait","P-value","-log10P","Trait description"),rownames=F)))
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
    }else{
    cat("  \n No information for other phenotypes at this SNP \n  ")
    cat('\\pagebreak  ')

    }
    }else{

    cat("  \n No pheWAS information \n  ")
    cat('\\pagebreak  ')
    }

cat("  \n")
cat("  \n### PheWAS: Lowest P-values for other phenotypes in a 3Mb window \n ",  qtls$trait_ref_name[i], "  \n")
if(nrow(table2)>0){

phewas_table2<-table2[which(table2$trait_ref_name %in% qtls$trait_ref_name[i]),]
if(nrow(phewas_table2)>0){
rownames(phewas_table2)<-NULL
phewas_table2$trait2_name<-gsub("_"," ",phewas_table2$trait2_name)
print(htmltools::tagList(datatable(phewas_table2[,c("dataset","trait2_name","trait2_topsnp","log10P","r2","dprime","trait_description")],colnames=c("Dataset","Trait","Trait topSNP","-log10P","r2","dprime","Trait description"),rownames=F,escape=T,options = list(pageLength = 10,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  ))))

cat("<br>")
cat("  \n")


if(nrow(phewas_table2)>4){
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
cat("<br>")
cat("  \n")
}



    }else{

cat("  \n No information for other phenotypes in 3Mb window of trait topSNP \n  ")
cat("<br>")
cat("  \n")

    }

    }else{

      cat("  \n No pheWAS information \n  ")
    cat('\\pagebreak  ')
    }



}









```
